---
title: "1_ICD9_Patient_Selection"
author: "Meg Hutch"
date: "February 27, 2020"
output: html_document
---

```{r}
library(dplyr)
```

```{r}
# All mimiciii icd9 codes/ids
icd <- read.csv("/Users/samuelhamilton/Downloads/Mimic_HSIP/Mimic_Data/diagnoses_icd.csv")

# diagnosis definitions
d_icd <- read.csv("/Users/samuelhamilton/Downloads/Mimic_HSIP/Mimic_Data/d_icd_diagnoses.csv")

# Import Admissions
admissions <- read.csv("/Users/samuelhamilton/Downloads/Mimic_HSIP/Mimic_Data/admissions.csv")

# remove row.ids
d_icd$row_id <- NULL
icd$row_id <- NULL
```

Add the diagnosis definitions to the icd9 codes
```{r}
icd <- merge(icd, d_icd, by = "icd9_code")
```

Query patients with Cardiac Arrest & Create an ID dataframe
```{r}
ca <- icd %>% filter(icd9_code == "4275")
nrow(unique(data.frame(ca$subject_id)))

ca_ids <- ca %>% select(subject_id, hadm_id) %>% distinct
```

For patients who might have multiple cardiac arrests, link the hadm_ids to the admissions table in order to identify the first cardiac icd9 code occurence

```{r}
ca_admissions <- merge(ca_ids, admissions, by = c("subject_id", "hadm_id"))

# convert admittime to posixct
ca_admissions$admittime_conv<- strptime(as.character(ca_admissions$admittime), format = "%Y-%m-%d %H:%M:%S", tz = "EST")

ca_admissions$admittime_conv <- as.POSIXct(ca_admissions$admittime_conv, tz="EST")

# group by subject_id and arrange by date
ca_admissions <- ca_admissions %>% group_by(subject_id) %>% arrange(admittime_conv) %>% slice(1L)
```

Modify the ca_ids df to contain the confirmed hospital admission with first icd9 code for ca
```{r}
ca_ids <- ca_admissions %>% select(subject_id, hadm_id) 
```

How many patients died -- nice, pretty balanced

Note: Hospital_expire_flag indicates whether the patient died within the given hospitalization. 1 indicates death in the hospital, and 0 indicates survival to hospital discharge."
```{r}
table(ca_admissions$hospital_expire_flag)
```

Add the morality label to the ids
```{r}
mortality <- ca_admissions %>% select(subject_id, hospital_expire_flag)

ca_ids <- merge(ca_ids, mortality, by = "subject_id")
```


Export final patient list for further database querying
```{r}
write.csv(ca_ids, file = "C:/Users/User/Box Sync/Projects/Mimic_HSIP/Mimic_Data/Cardiac_Arrest/ca_ids.csv", row.names = FALSE)
```
