---
title: "2_Physiologic_Data_HR"
author: "Meg Hutch"
date: "February 27, 2020"
output: html_document
---

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

```{r}
# Heart Rate Data
hr <- read.csv("C:/Users/User/Box Sync/Projects/Mimic_HSIP/Mimic_Data/Cardiac_Arrest/Chartevents_211_hr.csv")

# Patient IDs
ca_ids <- read.csv("C:/Users/User/Box Sync/Projects/Mimic_HSIP/Mimic_Data/Cardiac_Arrest/ca_ids.csv")
```

Patients with Heart Rate Data
```{r}
nrow(unique(data.frame(hr$subject_id)))
nrow(unique(data.frame(hr$hadm_id))) # checking that there are unique admissions per patient - looks good!
```

How many measurements per patient
```{r}
hr_counts <- hr %>% group_by(subject_id) %>% count(num_obs = n())

hr_counts_200 <- hr_counts %>% filter(num_obs >= 200)
```

We will need to check the time interval/perhaps standardize by time in some fashion (ie: if multiple measurements within the same time frame, do we want to take the mean, max, min? min or max heart rate can be bad)


## Data Pre-processing

Format the time into Posixct Format
```{r}
# CHF Patients
hr$charttime_conv <- strptime(as.character(hr$charttime), format = "%Y-%m-%d %H:%M:%S", tz = "EST")
hr$charttime_conv <- as.POSIXct(hr$charttime_conv, tz="EST")
```

Remove values if NULL
```{r}
hr <- hr %>% filter(!value == "NULL")
```

Convert values to numeric
```{r}
hr$value <- as.numeric(as.character(hr$value))
```


Take the min, max, mean, and sd when measurements occur within the same time interval (if we want this step!)
```{r}
#hr <- hr %>% group_by(subject_id, charttime_conv) %>% mutate(max_hr = max(value, na.rm = TRUE)) %>% mutate(min_hr = min(value, na.rm = TRUE)) %>% mutate(mean_hr = mean(value, na.rm = TRUE)) %>% mutate(sd_hr = sd(value, na.rm = TRUE)) 
```

## Create time plots
```{r}
# x = subject_id
plot_hr <- function(x) { 
  
  # arrange by Date
  df <- hr %>% filter(subject_id == x) %>% arrange(charttime_conv)

  #Title: create a vector that will contain the ID number and can be used to title the    graph by patient ID 
  title <- x
  
  #Create limits to identify the start and end dates of the time series
  first <- first(df$charttime_conv)
  last <- last(df$charttime_conv)

  hr <- df %>% filter(!is.na(value)) %>% select(subject_id, charttime_conv, value) %>%
    ggplot(aes(x = charttime_conv, y = value)) + 
    geom_point(color = "slateblue") + 
    geom_line(color = "slateblue") +
    #geom_hline(yintercept=3) + #could demark dangerous (low or high) HRs
    scale_y_continuous(name = "BMP") + #define the y-axis limits
    #scale_x_continuous(name = "Time", date_break = "1 hour", date_labels = "%H", limits = c(first, last)) + #, limits = c(-6, 0), breaks = c(-6:0)) +
    scale_x_datetime(name = "Time", date_break = "6 hour", date_labels = "%H", limits = c(first, last)) +
    ggtitle(paste0('ID:', title))  #add a title to identify patients 

  #remove legend and x-axis
  hr <- hr  + theme(legend.position = "none")

  #remove gray background and add a border around the plot
  hr <- hr + theme(panel.border = element_rect(fill=NA,color="black", size=0.5,
                                               linetype="solid"))
  
  
  hr <- hr + theme_bw()
  
  return(hr)
}

plot_hr(3) # not deceased
plot_hr(12) # deceased patient

```

