---
title: "Match_Waveform_IDs_ICD9"
author: "Meg Hutch"
date: "February 2, 2020"
output: html_document
---

```{r}
library(dplyr)
```

Import all mimic icd9 diagnoses
```{r}
# All mimiciii icd9 codes/ids
icd <- read.csv("C:/Users/User/Box Sync/Projects/Mimic_HSIP/Mimic_Data/diagnoses_icd.csv")

# mimic waveform ii ids
wf_ids <- read.table(file = "C:/Users/User/Box Sync/Projects/Mimic_HSIP/Processed_Data/mimicii_matched_records.txt", sep = ";",colClasses = 'character')

# change colnames - subject_id
colnames(wf_ids) <- c("subject_id", "subject_id_wf") 
```

Identify icd ids that overlap with the wf_ids
```{r}
wf_icd_ids <- merge(wf_ids, icd, by = "subject_id")

# keep only unique wf_icd_ids - the subject_id_wf is the timestamp of a unique waveform so for now, we do not need it!
wf_icd_ids$subject_id_wf <- NULL

# keep distinct rows
wf_icd_ids <- wf_icd_ids %>% distinct
```

We have 1,763 unique record ids
```{r}
nrow(unique(data.frame(wf_icd_ids$subject_id)))
```

# Count number of ICD
```{r}
wf_icd_ids <- wf_icd_ids %>% group_by(icd9_code) %>% mutate(count_icd = n()) %>% arrange(desc(count_icd))

# check unique icd9 counts
icd_counts <- wf_icd_ids %>% select(icd9_code, count_icd) %>% distinct
```

Most Frequent ICD-9 counts - checked definition from google
```{r}
# 428.0: Congestive heart failure, unspecified; 1,198 patients

# 401.9: Unspecified essential hypertension; 1,039 patients

# 414.01: Atherosclerotic heart disease of native coronary artery without angina pectoris; 907 patients

# 427.31: Atrial Fibrilation; 862

# 584.9: Acute kidney failure, unspecified; 600 Patients

# 250.00: Diabetes mellitus without mention of complication, type II or unspecified type, not stated as uncontrolled; 573 patients

```

Confirm definitions from the d_icd mimic table - they look to be matching my google search!
```{r}
d_icd <- read.csv("C:/Users/User/Box Sync/Projects/Mimic_HSIP/Mimic_Data/d_icd_diagnoses.csv")

# select only columns of interest
d_icd <- d_icd %>% select(icd9_code, short_title, long_title)

# match with icd_counts
icd_counts <- merge(d_icd, icd_counts, by = "icd9_code")

icd_counts <- icd_counts %>% arrange(desc(count_icd))
```

If we are interested in the NOS, Congestive heart patients, we can subset the waveform ids as such
```{r}
wf_icd_ids_chf <- wf_icd_ids %>% filter(icd9_code == 4280)


# Add back the wave form time stamps
wf_icd_ids_chf_chf <- merge(wf_icd_ids_chf, wf_ids, by = "subject_id")
```

export to csv
```{r}
write.csv(wf_icd_ids_chf, "C:/Users/User/Box Sync/Projects/Mimic_HSIP/Processed_Data/wf_icd_ids_chf.csv", row.names = FALSE)
```
